### 2022-08-05 chapter7/chapter.tex
ch 7/8 comments

%!TEX root = ../thesis.tex
\chapter{Conclusion\label{ch: conclusion}} 
A fluid-fluid queue is a stochastic fluid queue, where the driving process is a fluid queue itself. Fluid queues provide a model for a single continuous performance measure of a system in the presence of a random environment. Fluid queues have found a wide variety of applications including risk processes, telecommunications, and environmental modelling, among others. Given the success of fluid queues it is believed that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success. \cite{bo2014} provide an analysis of fluid-fluid queues and derive operator-analytic expressions for the first-return operator, and limiting distribution of a fluid-fluid queue.

@@

In Chapter~\ref{ch:galerkin} we introduced a discontinuous Galerkin (DG) scheme for the approximation of the generator of a fluid queue and show how to approximate the operators from \cite{bo2014}. High-order DG schemes are known to perform well for smooth problems, but problems can occur in the presence of discontinuities, such as negative probability estimates. In some contexts, slope limiting can be used to rectify this, but they result in approximations to operators which are non-linear which complicates the computation of the operators in \cite{bo2014}. The uniformisation scheme of \cite{bo2013}, which is equivalent to the simplest DG scheme, does not give erroneous approximations in the presence of discontinuities but can converge slowly compared to high-order DG schemes.

@@

In summary, this thesis investigated three main computational frameworks for the analysis of fluid queues and fluid-fluid queues, namely, the uniformisation, DG, and QBD-RAP schemes. The uniformisation scheme is from \cite{bo2013}. The DG method is well known, but its application to fluid and fluid-fluid queues has not been well-studied. This thesis introduces the DG scheme in the context of fluid and fluid-fluid queues. The QBD-RAP scheme is a novel approach which overcomes some known issues of the DG scheme, and converges faster than the uniformisation scheme. Analysis of the QBD-RAP scheme proves that it is convergent. Numerical experiments demonstrate the effectiveness of the approximation schemes for various problems.

>>>>

%!TEX root = ../thesis.tex
\chapter{Conclusion\label{ch: conclusion}} 
A fluid-fluid queue is a stochastic fluid queue, where the driving process is a fluid queue itself. Fluid queues provide a model for a single continuous performance measure of a system in the presence of a random environment. Fluid queues have found a wide variety of applications including risk processes, telecommunications, and environmental modelling, among others. Given the success of fluid queues it is plausible that the extension to fluid-fluid queues, which enable us to track two continuous performance measures of a system, will also find success. \cite{bo2014} provide an analysis of fluid-fluid queues and derive operator-analytic expressions for the first-return operator, and limiting distribution of a fluid-fluid queue.

@@

In Chapter~\ref{ch:galerkin} we introduced a discontinuous Galerkin (DG) scheme for the approximation of the generator of a fluid queue and show how to approximate the operators from \cite{bo2014}. High-order DG schemes are known to perform well for smooth problems, but problems, such as negative probability estimates, can occur in the presence of discontinuities. In some contexts, slope limiting can be used to rectify this, but they result in approximations to operators which are non-linear which complicates the computation of the operators in \cite{bo2014}. The uniformisation scheme of \cite{bo2013}, which is equivalent to the simplest DG scheme, does not give erroneous approximations in the presence of discontinuities but can converge slowly compared to high-order DG schemes.

@@

In summary, this thesis investigated three main computational frameworks for the analysis of fluid queues and fluid-fluid queues, namely, the uniformisation, DG, and QBD-RAP schemes. The uniformisation scheme is from \cite{bo2013}. The DG method is well known, but its application to fluid and fluid-fluid queues has not been well-studied. The QBD-RAP scheme is a novel approach which overcomes some known issues of the DG scheme, and converges faster than the uniformisation scheme. Analysis of the QBD-RAP scheme proves that it is convergent. Numerical experiments demonstrate the effectiveness of the approximation schemes for various problems.

### 2022-08-05 intro/chapter.tex
ch 7/8 comments

The addition of the second level process \(\{Y(t)\}\) greatly increases the flexibility and modelling potential. However, the added flexibility comes at a cost to the in the analysis of the model. The analysis of \cite{bo2014} is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). For practical computation of the results of \cite{bo2014}, a discretisation of the infinitesimal generator of the fluid queue can be used. To this end, to date, two possible discretisations have been suggested. Taking a differential equations-based approach, \cite{blnos2022} use the discontinuous Galerkin (DG) method to discretise the operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process. The QBD approximation of \cite{bo2013} is derived via a uniformisation argument, so we refer to it as the uniformisation scheme throughout this thesis. Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations. It turns out that the uniformisation scheme of \cite{bo2013} is a subclass of the former; the uniformisation scheme can be viewed as the simplest DG scheme where cells are all have equal width, the operator is projected onto a basis of piecewise constant functions, and an upwind flux is used to approximate the flow of mass between cells. This is also equivalent to a finite-volume scheme.

>>>>

The addition of the second level process \(\{Y(t)\}\) greatly increases the flexibility and modelling potential. However, the added flexibility comes at a cost to the in the analysis of the model. For both fluid-fluid queue and classical fluid queues, performance measures such has the stationary distribution, and the Laplace-transforms of hitting times and return times, can be derived in terms of the infinitesimal generator of the modulating process. For fluid queues the modulating process is a CTMC for which the infinitesimal generator is a matrix. Thus, expressions for the performance measures of fluid queues are matrix expressions which are readily computable. For fluid-fluid queue the modulating processes is a fluid queue for which the infinitesimal generator is a matrix of differential operators. Thus, expressions for the performance measures of fluid-fluid queues are matrices of differential operators which, in all but the simplest of cases, are not readily computable.

The analysis of \cite{bo2014} is in principle similar to the matrix-analytic methods of \cite{bean2005}, and derives results about the second level process \(\{Y(t)\}_{t\geq0}\) in terms of the infinitesimal generator (a differential operator) of the fluid queue, \(\{(X(t),\varphi(t))\}_{t\geq0}\). For practical computation of the results of \cite{bo2014}, a discretisation of the infinitesimal generator of the fluid queue can be used. To this end, to date, two possible discretisations have been suggested. Taking a differential equations-based approach, \cite{blnos2022} use the discontinuous Galerkin (DG) method to discretise the operator, while \cite{bo2013} take a stochastic modelling and matrix-analytic methods approach to approximate the fluid-queue by a quasi-birth-and-death (QBD) process. The QBD approximation of \cite{bo2013} is derived via a uniformisation argument, so we refer to it as the uniformisation scheme throughout this thesis. Both approaches are insightful and offer different tools and perspectives with which to analyse the resulting approximations. It turns out that the uniformisation scheme of \cite{bo2013} is a subclass of the former; the uniformisation scheme can be viewed as the simplest DG scheme where cells are all have equal width, the operator is projected onto a basis of piecewise constant functions, and an upwind flux is used to approximate the flow of mass between cells. This is also equivalent to a finite-volume scheme.
